---
title: "Report of Flight Price Predicton"
author: "071970225 Zhoukang"
date: "2022-11-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
if(!require(readr)) install.packages("readr")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(stringr)) install.packages("stringr")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(dslabs)) install.packages("dslabs")
if(!require(data.table)) install.packages("data.table")
if(!require(ggrepel)) install.packages("ggrepel")
if(!require(ggthemes)) install.packages("ggthemes")
if(!require(magrittr)) install.packages("magrittr")
if(!require(caret)) install.packages("caret")
if(!require(openxlsx)) install.packages("openxlsx")
if(!require(randomForest)) install.packages("randomForest")
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
library(gridExtra)
library(dslabs)
library(data.table) # for fread if required
library(ggrepel) # for some plots it modifies the position of labels 
library(ggthemes) # cool ggplot themes, check it out
library(tidyr)
library(tidyverse)# The pipe sign %>% is in this package
library(magrittr) #The pipe sign %<>% and %$% are in this package
library(lubridate)#The function "as_datetime"
library(caret)#The function "createDataPartition.
library(openxlsx) #The function read.xlsx
library(randomForest) 
```

```{r,include=FALSE}
#The xlsx file can be downloaded from the github repository.
#First we load the data file into R.
dl<-("./Data_Train.xlsx") #This is the relative path for R to find the file.
#Make sure you have put the xlsx file in your current working directory.
#Use "getwd()"to get your current working directory.
Raw<-read.xlsx(dl)
```


# 1 Introduction
## 1.1 The The information about Dataset "Flight Price Prediction"
|      The information below comes from the author of this dataset,you can check in the website:https://www.kaggle.com/datasets/jillanisofttech/flight-price-prediction-dataset. The objective of the study is to analyze the flight booking dataset obtained from the “Ease My Trip” website and to conduct various statistical hypothesis tests in order to get meaningful information from it. The 'Linear Regression statistical algorithm would be used to train the dataset and predict a continuous target variable. 'Easemytrip' is an internet platform for booking flight tickets, and hence a platform that potential passengers use to buy tickets. A thorough study of the data will aid in the discovery of valuable insights that will be of enormous value to passengers. 
## 1.2 The goal of the project
|      The goal of this project is to explore the relation between the flight price with other predictors such as airline,date,source,destination and duration.
|      As a machine learing project,Only the training set of this dataset will be used     because the testing set provided by the author did not include the price whereby no test can be implemented on it.  
|      In order to check the model precison,the training set will be de split into two part and one part will be remained as the testing set.

# 2 Methods and Analysis
## 2.1 Data cleaning
|      First the dataset should be cleaned to be fit for analyzing.We can see  the first several rows of the dataset:
```{r}
head(Raw)
```  

|      The general information of the dataset is as below:
```{r}
summary(Raw)
```  
|      We can see that some columns are character and they will be converted into factors and be marked with unique number.   
```{r,include=FALSE}
Data<-Raw%>%mutate(Airline=as.factor(Airline),Date=dmy(Date_of_Journey),
               Source=as.factor(Source),Destination=as.factor(Destination),
               Total_Stops=sapply(Raw$Total_Stops,function(TS) ifelse(TS=="non-stop",0,
              ifelse(TS=="1 stop",1,ifelse(TS=="2 stops",2,ifelse(TS=="3 stops",3,4))))))%>%
select(Airline,Date,Source,Destination,Duration,Total_Stops,Price)

#Then we convert the factor into numeric.
Data%<>%
mutate(Airline=sapply(Data$Airline, function(Airline) which(Airline==levels(Data$Airline))))%>%
mutate(Source=sapply(Data$Source,function(Sc) which(Sc==levels(Data$Source))))%>%
mutate(Destination=sapply(Data$Destination,function(Ds) which(Ds==levels(Data$Destination))))%>%
mutate(Duration_hour=as.numeric(str_extract(str_extract(Duration,"^\\d+[h]"),"\\d+"))+
ifelse(nchar(Duration)<=3,0,as.numeric(str_extract(str_extract(Duration,"\\d+[m]$"),"\\d+"))/60))%>%
select(-Duration)
```
|      After this work we can see a cleaner dataset:
```{r}
head(Data)
```  
|      The we check the basic information of the dataset again:
```{r}
summary(Data)
```  
|      Now We see one NA in Duration and Total_Stops which shows there are some unnormal values in the dataset.Have a glimpse at it:
```{r}
na1<-which(is.na(Data$Duration))
na2<-which(is.na(Data$Total_Stops))
#Check the unnormal value in the Raw dataset
Raw[na1,]
Raw[na2,]
```  
|      The NA in the Total_Stops is caused by missing value.The time for a flight is impossible to be 5 minutes so we will remove these two rows.
```{r,include=FALSE}
index_delete<-c(na1,na2)
Data_cor<-Data%>%slice(-index_delete)
```  
|      Now we have finished the data cleaning.Then we need to split the dataset into training set and Testing set.The testing set will be 30% of the whole dataset.
```{r,include=FALSE}
set.seed(1,sample.kind = "Rounding")
test_index<-createDataPartition(Data_cor$Price,times = 1,p = 0.3,list = FALSE)
Train_set<-Data_cor%>%slice(-test_index)
Test_set<-Data_cor%>%slice(test_index)
```  

## 2.2 Visualization
|      Visualization is a good way to search for some relations between the price and other predictors.

### 2.2.1 The Airline and Price
```{r,echo=FALSE}
Train_set%>%boxplot(Price~Airline,data = .)
```

### 2.2.2 The Date and Price
```{r,echo=FALSE}
Train_set%>%boxplot(Price~Date,data = .)
```

### 2.2.3 The Source and Price
```{r,echo=FALSE}
Train_set%>%boxplot(Price~Source,data = .)
```

### 2.2.4 The destination and Price
```{r,echo=FALSE}
Train_set%>%boxplot(Price~Destination,data = .)
```

### 2.2.5 The Total_Stops and Price
```{r,echo=FALSE}
Train_set%>%boxplot(Price~Total_Stops,data = .)
```

### 2.2.6 The Duration_hour and Price
```{r,echo=FALSE,message=FALSE}
Train_set%>%ggplot(aes(Duration_hour,Price))+geom_point()+geom_smooth()
```

### 2.2.7 Summary of the predictors
|      To  be frank,not any strong relation is showed between the price and other predictors. The analysis be based on all these predictors.

# 3 The build of the model 
## 3.1 The prediction of price
|      In order to predict the price,other 6 predictors will be considered and be used in 3 different algorithm which are glm(General liner model),knn(k-Nearest Neighbor) and rf(Random Forest).  
|      For the estimate of the model,we build the RMSE to compute the distance between the estimated price and the true price.    
|      To compute the avarage of all price: 
```{r,include=TRUE,echo=TRUE}
mu<-mean(Train_set$Price)
```  
|      Now we can check the RMSE for the first time to see our precision.Define the function RMSE as below:
```{r,include=TRUE,echo=TRUE}
RMSE<-function(pred_value,true_value)
{sqrt(mean((pred_value-true_value)^2))
}
```  
|      And the RMSE for the basic ratings and true ratings is listed:
```{r,include=TRUE,echo=TRUE}
RMSE(mu,Train_set$Price)
```  
## 3.2 The knn model
```{r}
model_knn<-train(Price~.,method="knn",data =Data_cor,TuneGrid=data.frame(k=seq(1,20,1)))
pred_knn<-predict(model_knn,Test_set)
```
We can see the RMSE for knn model is: 
```{r}
RMSE(pred_knn,Test_set$Price)
```  
## 3.3 The glm model 
```{r}
model_glm<-train(Price~.,method="glm",data = Data_cor)
pred_glm<-predict(model_glm,Test_set)
```
We can see the RMSE for glm model is:
```{r}
RMSE(pred_glm,Test_set$Price)
```    
## 3.4 The rf model
```{r}
set.seed(123,sample.kind = "Rounding")
model_rf<-randomForest(Price ~ ., data = Train_set, importance = TRUE)
pred_rf<-predict(model_rf,Test_set)
```
We can see the RMSE for glm model is:
```{r}
RMSE(pred_rf,Test_set$Price)
```  
## 3.5 The ensemble of 3 models
|      We simply compute the average of every model because we can not use the result generated from the testing set.
```{r}
pred_ensemble<-(pred_knn+pred_glm+pred_rf)/3
```  
|      And we can check the RMSE for the testing dataset:
```{r}
RMSE(pred_ensemble,Test_set$Price)
```  
|      We can compare our prediction with just guessing the average price for all price in the testing dataset.
```{r}
mu_test<-mean(Test_set$Price)
RMSE(mu_test,Test_set$Price)
```
|      Now we can see our model do have some improvements in predicting the price of one flight though not very much.  
# 4 The results and conclusion
|      Though 3 different model are employed and ensemble to predict the price, the result was not satisfying and the RMSE was still quite huge.From the visualization we can not see a obvious relation between the price and other predictors.In normal sense,a longer duration means a longer distance and the price for that flight would be higher while in this dataset it seems not to be in this case.































































































































